HO unification from object language to meta language

### 2024-02-21 paper/main.tex
minor changes
update tex

\section{Introduction}

@@

\noindent
The encoding of DTT provided by Elpi, that we will discuss at length later in
\cref{sec:encodings, ...} features the following term constructors:

@@

\section{Lanuages description}

@@

In order to reason about unification of the terms of an objet language within a
meta language, we start by formally describing the two languages. We use
meta-programming to accomplish this result. In \cref{code:lang-descr}, we
provide on the left the typing structure of the terms of the object language
together with the signature of the \eqfo function.

In this encoding, the fo\_tm type stands for the first order representation of
the object language terms. We model unification variables as integers
corresponding to memory addresses. This memory is represented with the subst
type abbreviation which is list of optional terms. If the cell $i$ is none, then
the variable $i$ is not assigned.

% Why the fo prefix for fo_tm && Uvar in OL have no scope
Because of our translation of the terms of the object language into the meta
language we have sketched in section 1, a object language term has no scope.
An example of this claim can be given by analyzing deeply the subterm \coqIn{P x} from
the instance forall\_dec. \coqIn{P} is an higher order variables
with type \coqIn{A ~$\to$~ Prop} and
\coqIn{x} is a name bound to \coqIn{P}.
However, at the meta level, the translation of \coqIn{P x} become \elpiIn{app[uv
1, x]}, that is, \elpiIn{uv 1} cannot see \elpiIn{x}.

Since our goal is to work with higher order variables, we need to compile the
term received in entry into a more expressive version. This is possible by using
a second type structure for the terms of the OL. This term definition is given
in the second column of the \cref{code:lang-descr}. The main different here is
that a unification variable (always identified with by an integer) has now a
list of terms representing the scope of the variable. We symbolize the
abstractions of the ML using the constructor \elpiIn{abs} of the type
\elpiIn{assmt}. The constructor \elpiIn{val} is is used to contain terms of the
object language. In this context we do not need to have an application node for
the ML.

>>>>

\section{Introduction}
\label{sec:intro}

@@

\noindent
The encoding of DTT provided by Elpi, that we will discuss at length later in
\cref{sec:encodings,sec:lang-spec}, features the following term constructors:

@@

\section{Languages description}
\label{sec:lang-spec}

@@

In order to reason about unification of the terms of an objet language within a
meta language, we start by formally describing the two languages. Employing
meta-programming for this purpose, \cref{code:lang-descr} presents, on the left,
the typing structure of object language terms along with the signature of the
\eqfo function.

In this encoding, the fo\_tm type stands for the first order representation of
the object language terms. We model unification variables as integers
corresponding to memory addresses. The memory is represented as list of optional
terms (the subst\_fo type abbreviation). If the cell $i$ is none, then the
variable $i$ is not instantiated, otherwise it has already been assigned to some
value.

% Why the fo prefix for fo_tm && Uvar in OL have no scope
Because of our translation of the terms of the object language into the meta
language outlined in \cref{sec:intro}, the variables of the object language have 
no scope.
An illustration of this observation can be found by analyzing deeply the subterm \coqIn{P x} from
the instance forall\_dec. Here, \coqIn{P} is an higher order variables
with type \coqIn{A ~$\to$~ Prop} and
\coqIn{x} is a name bound to \coqIn{P}.
However, at the meta level, the translation of \coqIn{P x} becomes \elpiIn{app[uv
1, x]}, that is, \elpiIn{uv 1} cannot reference \elpiIn{x}.

As our objective is to handle higher-order variables, a viable solution involves
addressing this issue through the compilation of the term received as input into
a more expressive version. This can be achieved by employing a second type
structure for the terms of the OL, as outlined in the second column of
\cref{code:lang-descr}. The primary difference lies in the fact that a
unification variable (always identified by an integer) now takes a list of terms
representing its scope. The abstractions of the ML are
symbolized using the constructor \elpiIn{abs} of the type \elpiIn{assmt}. The
constructor \elpiIn{val} is utilized to contain terms of the object language,
and in this context, there is no need for an application node in the ML.
